---
title: "Legionella delay"
author: "Dan Weinberger"
date: '2022-08-03'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(dplyr)
library(reshape2)
library(rjags)
```

## Background
The goal for this analysis is to reconstruct a time series of legionella infections using a time series based on symptom onset date and a delay distribution. This latent time series will then be linked to climate variables. This will be accomplished in a Bayesian framework using rjags.



### First, generate a time series based on exposure date..we will then apply a delay distribution to this to get an 'observed' time series, and see if we can then

```{r}
set.seed(123)
df1  <- cbind.data.frame('dates'=seq.Date(from=as.Date('2012-01-01'), to=as.Date('2022-07-01'), by='day'))

df1 <- df1 %>%
  mutate(t=row_number(), 
         sin1= 2*sin(2*pi*t/365),
         N_exposures = rpois(length(dates), exp(1 + 0.3*sin1 + rnorm(length(dates),0,0.75))) ) 

plot(df1$N_exposures)
```

Distribution to estimate the time between exposure and symptom onset. It's bounded at 20 days which is about the uppermost end.
```{r}
x <- seq(0, 20, by=1)
delay_dist = dgamma(x, shape = 4.96, scale = 1.27, log = FALSE)
plot(delay_dist, type = 'l')

#rmultinom(1,100,delay_dist)
```
Take random draws from a multinomial, based on N exosures on that day and the delay distribution to determine how many cases that were exposed on date t have symptom onset with a delay of d, with d{1:21}

```{r}
set.seed(1234)
dist_data <- t(sapply(df1$N_exposures, function(x) rmultinom(1,x,delay_dist)))

df2 <- cbind.data.frame('dates'=df1$dates, dist_data)

# Reshape the date into long format, determine the symptom onset date based on exposure date + delay, then sum by symptom onset date to get 'observed' time series

observed_ts <- melt(df2, id.vars='dates') %>%
  rename(exposure_date=dates) %>%
  mutate(delay = as.numeric(as.character(variable)),
         symptom_date= exposure_date + delay) %>%
  group_by(symptom_date) %>%
  summarize(N_cases=sum(value))
```

Compare the time series of exposure and time series of cases
```{r, fig.width=6, fig.height=3}
par(mfrow=c(1,2))

plot(df1$N_exposures, type='l', main='N exposures', bty='l')
  
plot(observed_ts$N_cases, type='l', main="N cases based on sympom onset", bty='l')

```

OK, now we can try to go backwards, let's reconstruct the exposure time series based on the case time series

```{r}
#lambda is the mean of the observed cases, based on symptom onset
#Delta is the reporting triangle that links observed time series (lambda) and latent time series (epsilon)
#

model_string <- "
model{
for(t in 1:N_times){
  N_cases[t] ~ dpois(lambda[t])

for(d in (1:21)){
  lambda[t] <- sum(delta[t,d]) #delta has dimensions [t,d]

  #These are equivalent, neither works.
  #epsilon[t] <- delta[t,d] / delay_dist[d] 
  delta[t,d] <- epsilon[t] * delay_dist[d] 

  }
 }
}
"



##############################################################
#Model Fitting
##############################################################
inits1=list(".RNG.seed"=c(123), ".RNG.name"='base::Wichmann-Hill')
inits2=list(".RNG.seed"=c(456), ".RNG.name"='base::Wichmann-Hill')
inits3=list(".RNG.seed"=c(789), ".RNG.name"='base::Wichmann-Hill')


##############################################
#Model Organization
##############################################
model_spec<-textConnection(model_string)
model_jags<-jags.model(model_spec, 
                       inits=list(inits1,inits2, inits3),
                       data=list('N_cases'=observed_ts$N_cases,
                                 'N_times'=nrow(observed_ts),
                                                       'delay_dist'=delay_dist
                                    ),
                       n.adapt=10000, 
                       n.chains=3)


```



