---
title: "JAGS_IncubationPeriod_Dec2023"
output: html_document
date: "2023-12-19"
---
Last update: 13 Feb 2024


Import Libraries
```{r setup, include=FALSE}

library(lubridate)
library(dplyr)
library(reshape2)
library(rjags)
library(stringr)
library(gtools)
library(HDInterval)
library(ggplot2)
library(coda)
library(fitdistrplus)


```



Delay Dist 
From a point source outbreak defined in Bays et al. 
```{r}
x <- seq(0, 20, by=1)
delay_dist = dgamma(x, shape = 4.96, scale = 1.27, log = FALSE)

```


Delay Dist from NC DHHS
```{r}

NCDHHS = read.csv("Documents/IncubationPeriod_Analysis/NC DHHS Data/DUA_LG_FairOutbreak_Linelist_11NOV2023_Clean.csv")

NCDHHS = NCDHHS[NCDHHS$Employee != 1,]
NCDHHS = NCDHHS[!is.na(NCDHHS$Visit.Date.1),]
NCDHHS$visit1 = as.Date(NCDHHS$Visit.Date.1, format = '%m/%d/%y')
NCDHHS$symptom.date = as.Date(NCDHHS$Symptom.Onset.Date, format = '%m/%d/%y')
NCDHHS = NCDHHS[!is.na(NCDHHS$symptom.date),]

NCDHHS$incubation.period = NCDHHS$symptom.date - NCDHHS$visit1
NCDHHS$case1 = 1 
NCDHHS1 = NCDHHS[!(NCDHHS$incubation.period < 0),]
NCDHHS1 = NCDHHS1 %>%
            filter(!is.na(NCDHHS1$incubation.period)) # 95 
NCDHHS1  = NCDHHS1 %>% 
          filter(Employee != 1 & incubation.period > 0)

# first day of visit only 
fit.gamma <- fitdist(as.numeric(NCDHHS1$incubation.period), "gamma", method = "mme")
summary(fit.gamma)
x.x <- seq(0, 20, by=1)
delay_dist1 = dgamma(x.x, shape = 5.3590725, scale = 1/0.7300765, log = F)

 
#---------------------------------------------------------------------------------------#
# for cases with multiple dates of visit, randomly select first or last date of visit for them: 


NCDHHS.short = NCDHHS[NCDHHS$Visit.Date.2 != '',c('Event.ID','Visit.Date.1','Visit.Date.2')]
facet = tidyr::gather(NCDHHS.short, group, visit.date, Visit.Date.1:Visit.Date.2)
set.seed(1)
facet1 = facet %>% 
  group_by(Event.ID) %>% 
  sample_n(1)
facet1$visit1 = as.Date(facet1$visit.date, format = '%m/%d/%y')
facet1 = facet1 %>%
          dplyr::select(Event.ID, visit1)

NCDHHS.long = NCDHHS[NCDHHS$Visit.Date.2 == '',c('Event.ID','Visit.Date.1')]
NCDHHS.long$visit1 = as.Date(NCDHHS.long$Visit.Date.1, format = '%m/%d/%y')
NCDHHS.long = NCDHHS.long %>%
          dplyr::select(Event.ID, visit1)

NCDHHS.all = rbind.data.frame(NCDHHS.long, facet1) %>%
              filter(!is.na(visit1))

NCDHHS.all1 = merge(NCDHHS.all, NCDHHS[,c('Event.ID', 'Symptom.Onset.Date')], by = 'Event.ID', all.x = T)

NCDHHS.all1$symptom.date = as.Date(NCDHHS.all1$Symptom.Onset.Date, format = '%m/%d/%y')
NCDHHS.all1$incubation.period = NCDHHS.all1$symptom.date - NCDHHS.all1$visit1

NCDHHS.all1$case1 = 1
NCDHHS.all1 = NCDHHS.all1[!(NCDHHS.all1$incubation.period < 0),]


fit.gamma <- fitdist(NCDHHS.all1incubation.period, "gamma", method = "mme")
summary(fit.gamma)
x.x <- seq(0, 20, by=1)
delay_dist.ncdhhs = dgamma(x.x, shape = 5.1225291, scale = 1/0.7362183, log = F)


```



CT DPH Case Import Jan 2011 to Dec 2021
```{r}
#import travel/HCA data 

df  =  case.df  %>%
            filter(DISEASE_STATUS.x != 'REVOKED') %>%
            filter(STATE_HIDDEN %in% c('CT','')) %>%
            filter(!TYPE_OF_LEGIONELLOSIS.x %in% c("PONTIAC FEVER (FEVER, MYALGIA WITHOUT PNEUMONIA)", 
                                                   "Pontiac Fever (fever, myalgia without pneumonia)",
                                                   "OTHR")) %>%
            filter(TRAVEL_BEFORE_SYMPTOM_ONSET != 'Yes') %>%
            filter(LEG_TRVL_CITY_0 == "") %>% # 921
            filter(EVENT_TYPE.y == 'SYMPTOM_ONSET_DATE') # %>% #737
           # filter(NOSOCOMIAL != 'PRESUMPT_HCA') # 714


df$case1 = 1
study.period<- cbind.data.frame('dates'=seq.Date(from=as.Date('2011-01-01'), to=as.Date('2021-12-31'), by='day'))


case.df.1 = setNames(aggregate(df$case1 ~ df$ddate.y, FUN = 'sum'), c("ddate", "case_sum"))
case.df = merge(study.period, case.df.1, by.x = c('dates'), by.y = c('ddate'), all.x = T)
case.df$case_sum = ifelse(is.na(case.df$case_sum), 0, case.df$case_sum)
 
```

 

Weather import 
- This data was cleaned and aggregated using the file: 

```{r}
weather.d = read.csv("Documents/IncubationPeriod_Analysis/Rainfall CSVs/weather_agg_humidity_state.csv")
weather.d$ddate = as.Date(as.character(weather.d$date1))
weather.d = weather.d[year(weather.d$ddate) != 2022,]
```

Select weather variable 
```{r}

weather.d$rain0 = weather.d$Reported_PrecipSum
weather.d = weather.d[,c('ddate','rain0')]


# or run analysis on relative humidity 
weather.d$rain0 = weather.d$Hourly_RelHum_Mean

# interaction between rel hum and rainfall 
quantile(weather.d$Hourly_RelHum_Mean, probs= c(0,1,0.75))
weather.d$hum_high = ifelse(weather.d$Hourly_RelHum_Mean > 81.81929, 1, 0) # 
weather.d$hum_low = ifelse(weather.d$hum_high == 1,0,1)
weather.d$rain0 = weather.d$Reported_PrecipSum
weather.d = weather.d[,c('ddate','rain0','hum_low','hum_high')]


```


Standardize weather

```{r}

test = weather.d
library(dplyr)
test0 = test %>% 
  group_by(year(ddate)) %>% 
  mutate(
    rain0_scale = as.numeric(scale(rain0)) 
    )
test1 = test0[,c('ddate','rain0_scale')]
colnames(test1) = c('ddate','rain0')
weather.d = test1
weather.d$ddate = as.Date(weather.d$ddate)
  
```



Randomize the rainfall data within each month and year 
```{r}


set.seed(123) # random 1
set.seed(456) # random 2
set.seed(789) # random 3 

weather.d = weather.d %>% 
  mutate(monthdate = paste(month(ddate),year(ddate))) %>%
  group_by(monthdate) %>% 
  mutate(random_rain0=sample(rain0))

saveRDS(
  weather.d,'Documents/IncubationPeriod_Analysis/legionella_delay/Results/rainfall_20Mar24_random3.rds')



```


Make observed TS with sin and cos vars 
```{r}

observed_ts = merge(case.df,weather.d, by.y = 'ddate', by.x = 'dates', all.x = T)
observed_ts$N_cases = observed_ts$case_sum

observed_ts$t = 1:nrow(observed_ts)

observed_ts$sin1 = sin(2*pi*observed_ts$t/365)
observed_ts$cos1 = cos(2*pi*observed_ts$t/365)

observed_ts$sin2 = sin(2*pi*observed_ts$t/182)
observed_ts$cos2 = cos(2*pi*observed_ts$t/182)

observed_ts$year = year(observed_ts$dates)


df = observed_ts

```


Base model that uses the previously published delay distribution (delay_dist) which extended 20 days 

## BASE Model 
```{r}

 sink("leadlag_AR1_20delay.txt")

 cat(
  model_string <-
" model{

for(t in 37:N_times){

N_cases[t] ~ dpois(y[t])

 y[t]  <- (z[t]*delay_dist[1] + 
               z[t- 1 ]*delay_dist[2] + 
               z[t- 2 ]*delay_dist[3] +
               z[t- 3 ]*delay_dist[4] + 
               z[t- 4 ]*delay_dist[5] +
               z[t- 5 ]*delay_dist[6] + 
               z[t- 6 ]*delay_dist[7] +
               z[t- 7 ]*delay_dist[8] + 
               z[t- 8 ]*delay_dist[9] +
               z[t- 9 ]*delay_dist[10] + 
               z[t- 10 ]*delay_dist[11] +
               z[t- 11 ]*delay_dist[12] + 
               z[t- 12 ]*delay_dist[13] +
               z[t- 13 ]*delay_dist[14] + 
               z[t- 14 ]*delay_dist[15] +
               z[t- 15 ]*delay_dist[16] + 
               z[t- 16 ]*delay_dist[17] +
               z[t- 17 ]*delay_dist[18] +
               z[t- 18 ]*delay_dist[19] +
               z[t- 19 ]*delay_dist[20] +
               z[t- 20 ]*delay_dist[21] )

}

for(t in 16:N_times){

z[t] ~ dpois(lambda[t])

log(lambda[t]) <- ( 

        int + rain[t-14]*beta[1] + rain[t-13]*beta[2] + rain[t-12]*beta[3] +  rain[t-11]*beta[4]  +   rain[t-10]*beta[5]
        + rain[t-9]*beta[6] + rain[t-8]*beta[7] + rain[t-7]*beta[8] +  rain[t-6]*beta[9] 
        + rain[t-5]*beta[10] + rain[t-4]*beta[11] + rain[t-3]*beta[12] +  rain[t-2]*beta[13]  +   rain[t-1]*beta[14]
        + rain[t]*beta[15] + rain[t+1]*beta[16] + rain[t+2]*beta[17] 
        +  rain[t+3]*beta[18] +  rain[t+4]*beta[19] +  rain[t+5]*beta[20] +  rain[t+6]*beta[21] +  rain[t+7]*beta[22]
        +  rain[t+8]*beta[23] +  rain[t+9]*beta[24] + rain[t+10]*beta[25] +  rain[t+11]*beta[26] + rain[t+12]*beta[27]
        + rain[t+13]*beta[28] +  rain[t+14]*beta[29]  
        + delta[1]*sin1[t] + delta[2]*cos1[t] + delta[3]*sin2[t] + delta[4]*cos2[t]
  + theta[1]*year[t]
  
        )
}

   int ~ dnorm(0, 1e-4)
 
 
beta[1] ~ dnorm(0.00,(sigma2_beta_inv*(1.00 - alpha^2)))
for(t in 2:29){
   beta[t] ~ dnorm((alpha*beta[t-1]),sigma2_beta_inv)
   }
sigma2_beta_inv ~ dgamma(0.01,0.01)
alpha ~ dunif(0,1.00)



  for(k in 1:4){
    delta[k] ~ dnorm(0, 1e-4)
   }
 
   

theta[1] ~ dnorm(0, 1e-4)

}
"
,fill = TRUE)
sink()

```


Adding an interaction for high/low humidity in the model 
```{r}

 sink("test_AR1_DLM_interaction.txt")

 cat(
  model_string <-
" model{

for(t in 37:N_times){

N_cases[t] ~ dpois(y[t])

 y[t]  <- (z[t]*delay_dist[1] + 
               z[t- 1 ]*delay_dist[2] + 
               z[t- 2 ]*delay_dist[3] +
               z[t- 3 ]*delay_dist[4] + 
               z[t- 4 ]*delay_dist[5] +
               z[t- 5 ]*delay_dist[6] + 
               z[t- 6 ]*delay_dist[7] +
               z[t- 7 ]*delay_dist[8] + 
               z[t- 8 ]*delay_dist[9] +
               z[t- 9 ]*delay_dist[10] + 
               z[t- 10 ]*delay_dist[11] +
               z[t- 11 ]*delay_dist[12] + 
               z[t- 12 ]*delay_dist[13] +
               z[t- 13 ]*delay_dist[14] + 
               z[t- 14 ]*delay_dist[15] +
               z[t- 15 ]*delay_dist[16] + 
               z[t- 16 ]*delay_dist[17] +
               z[t- 17 ]*delay_dist[18] +
               z[t- 18 ]*delay_dist[19] +
               z[t- 19 ]*delay_dist[20] +
               z[t- 20 ]*delay_dist[21] )

}

for(t in 16:N_times){

z[t] ~ dpois(lambda[t])

log(lambda[t]) <- ( 


        int + rain[t-14]*beta[1]*hum_high[t-14] + rain[t-13]*beta[2]*hum_high[t-13] + 
        rain[t-12]*beta[3]*hum_high[t-12] +  rain[t-11]*beta[4]*hum_high[t-11]  +   rain[t-10]*beta[5]*hum_high[t-10]
        + rain[t-9]*beta[6]*hum_high[t-9] + rain[t-8]*beta[7]*hum_high[t-8] + rain[t-7]*beta[8]*hum_high[t-7] +  
        rain[t-6]*beta[9]*hum_high[t-6] 
        + rain[t-5]*beta[10]*hum_high[t-5] + rain[t-4]*beta[11]*hum_high[t-4] + 
        rain[t-3]*beta[12]*hum_high[t-3] +  rain[t-2]*beta[13]*hum_high[t-2]  +   rain[t-1]*beta[14]*hum_high[t-1]
        + rain[t]*beta[15]*hum_high[t] + rain[t+1]*beta[16]*hum_high[t+1] + rain[t+2]*beta[17]*hum_high[t+2] 
        +  rain[t+3]*beta[18]*hum_high[t+3] +  rain[t+4]*beta[19]*hum_high[t+4] +  
        rain[t+5]*beta[20]*hum_high[t+5] +  rain[t+6]*beta[21]*hum_high[t+6] +  rain[t+7]*beta[22]*hum_high[t+7]
        +  rain[t+8]*beta[23]*hum_high[t+8] +  rain[t+9]*beta[24]*hum_high[t+9] + rain[t+10]*beta[25]*hum_high[t+10] + 
        rain[t+11]*beta[26]*hum_high[t+11] + rain[t+12]*beta[27]*hum_high[t+12]
        + rain[t+13]*beta[28]*hum_high[t+13] +  rain[t+14]*beta[29]*hum_high[t+14]  +
        
        
        rain[t-14]*beta[30]*hum_low[t-14] + rain[t-13]*beta[31]*hum_low[t-13] + 
        rain[t-12]*beta[32]*hum_low[t-12] +  rain[t-11]*beta[33]*hum_low[t-11]  +   rain[t-10]*beta[34]*hum_low[t-10]
        + rain[t-9]*beta[35]*hum_low[t-9] + rain[t-8]*beta[36]*hum_low[t-8] + rain[t-7]*beta[37]*hum_low[t-7] +  
        rain[t-6]*beta[38]*hum_low[t-6] 
        + rain[t-5]*beta[39]*hum_low[t-5] + rain[t-4]*beta[40]*hum_low[t-4] + 
        rain[t-3]*beta[41]*hum_low[t-3] +  rain[t-2]*beta[42]*hum_low[t-2]  +   rain[t-1]*beta[43]*hum_low[t-1]
        + rain[t]*beta[44]*hum_low[t] + rain[t+1]*beta[45]*hum_low[t+1] + rain[t+2]*beta[46]*hum_low[t+2]
        +  rain[t+3]*beta[47]*hum_low[t+3] +  rain[t+4]*beta[48]*hum_low[t+4] +  
        rain[t+5]*beta[49]*hum_low[t+5] +  rain[t+6]*beta[50]*hum_low[t+6] +  rain[t+7]*beta[51]*hum_low[t+7]
        +  rain[t+8]*beta[52]*hum_low[t+8] +  rain[t+9]*beta[53]*hum_low[t+9] + rain[t+10]*beta[54]*hum_low[t+10] +  
        rain[t+11]*beta[55]*hum_low[t+11] + rain[t+12]*beta[56]*hum_low[t+12]
        + rain[t+13]*beta[57]*hum_low[t+13] +  rain[t+14]*beta[58]*hum_low[t+14]
        
        
        + delta[1]*sin1[t] + delta[2]*cos1[t] + delta[3]*sin2[t] + delta[4]*cos2[t]
  + theta[1]*year[t]
  
        )
}

   int ~ dnorm(0, 1e-4)
 
 
beta[1] ~ dnorm(0.00,(sigma2_beta_inv*(1.00 - alpha^2)))
for(t in 2:29){
   beta[t] ~ dnorm((alpha*beta[t-1]),sigma2_beta_inv)
}
   
beta[30] ~ dnorm(0.00,(sigma2_beta_inv*(1.00 - alpha^2)))
for(t in 31:58){
   beta[t] ~ dnorm((alpha*beta[t-1]),sigma2_beta_inv)
}
   
sigma2_beta_inv ~ dgamma(0.01,0.01)
alpha ~ dunif(0,1.00)



  for(k in 1:4){
    delta[k] ~ dnorm(0, 1e-4)
   }
 
   

theta[1] ~ dnorm(0, 1e-4)

}
"
,fill = TRUE)
sink()


```


## RUN & STORE MODEL 

```{r}

##############################################################
#Model Fitting
##############################################################
inits1=list(".RNG.seed"= c(123), ".RNG.name"='base::Wichmann-Hill')
 inits2=list(".RNG.seed"= c(456), ".RNG.name"='base::Wichmann-Hill')
##############################################
#Model Organization
##############################################
 
# could give starting values - use values from model as starting point  

model_jags<-jags.model( 
  
                    # model_spec, 
                  # 'test_AR1_DLM_interaction.txt', 
                   'leadlag_AR1_20delay.txt',  
                       inits= list(inits1, inits2),
                       data=list('N_cases'=df$N_cases,
                                   'N_times'= 4000, 
                                 'delay_dist'=delay_dist,
                                 'sin1' = df$sin1,
                                 'cos1' = df$cos1,
                                 'sin2' = df$sin2,
                                 'cos2' = df$cos2,
                                 'year' = df$year,
                                 'rain'=df$rain0 # this changes depending on what the predictor is
                                # 'hum_high' = df$hum_high,
                                 # 'hum_low' = df$hum_low,
                                    ), 
                       n.adapt=5000,
                        n.chains=2) 
 

# parameters to monitor  
params<-c('beta', 'z', 'y', 'int', 'delta', 'alpha')


posterior_samples<-coda.samples(model_jags, 
                                params,  
                                n.iter=20000)
```



# Calc fit stats



Test the Geweke's convergence criteria

```{r}

beta_samples <- posterior_samples.all[,grepl('beta',colnames(posterior_samples.all))]
mcmc1 = as.mcmc(beta_samples) # need to isolate each chain and each beta 
geweke.plot(mcmc1)


```










